---
format: html
---

## 1. Hours Studied vs Performance Index

```{r}
#| echo: false
#| fig-width: 7
#| fig-height: 5
#| warning: false
#| message: false

library(tidyverse)

# Load actual dataset
df <- read_csv("StudentPerformance.csv")  # make sure the CSV is in the project folder

# Optional: replace spaces with underscores for easier plotting
names(df) <- gsub(" ", "_", names(df))

# Scatter plot
p1 <- ggplot(df, aes(x = Hours_Studied, y = Performance_Index)) +
  geom_point(aes(color = Previous_Scores, size = Sleep_Hours), alpha=0.8) +
  scale_color_gradient(low="#f4e1c1", high="#b35e0d") +
  theme_minimal(base_family="Georgia") +
  labs(
    title="Hours Studied vs Performance Index",
    x="Hours Studied",
    y="Performance Index",
    color="Previous Scores",
    size="Sleep Hours"
  ) +
  theme(
    plot.title = element_text(size=10, face="bold", hjust = 0.5),
    axis.title = element_text(size=9,hjust = 0.5),
    axis.text = element_text(size=8,hjust = 0.5)
  )

print(p1)
```

Students who study more hours generally achieve a higher Performance Index, showing a positive correlation.

Previous scores also influence performance - students with higher prior scores tend to cluster at higher Performance Index values.

Sleep hours affect performance subtly; students with very low or very high sleep tend to deviate slightly from the trend, suggesting balanced rest may help performance.

## 2. No.of Sample papers vs Performance Index

```{r}
#| echo: false
#| fig-width: 7
#| fig-height: 5
#| warning: false
#| message: false

# Aggregate Performance Index by number of sample papers

df_avg <- df %>%
group_by(Sample_Question_Papers_Practiced) %>%
summarise(Avg_Performance = mean(Performance_Index, na.rm=TRUE))

# Bar chart

p2 <- ggplot(df_avg, aes(x = factor(Sample_Question_Papers_Practiced), y = Avg_Performance, fill = Avg_Performance)) +
geom_bar(stat="identity") +
scale_fill_gradient(low="#f4e1c1", high="#b35e0d") +
theme_minimal(base_family="Georgia") +
labs(
title="Average Performance Index by Number of Sample Question Papers Practiced",
x="Number of Papers Practiced",
y="Average Performance Index"
) +
theme(
plot.title = element_text(size=10, face="bold",hjust = 0.5),
axis.title = element_text(size=9,hjust = 0.5),
axis.text = element_text(size=8,hjust = 0.5),
legend.position = "none"
)

print(p2)
```

Students who practice more sample question papers achieve higher average Performance Index, showing active preparation pays off.

The improvement is more noticeable when moving from 1-2 papers to 3-4 papers, suggesting a threshold effect where consistent practice significantly boosts performance.

Helps educators identify that encouraging practice with sample papers can be an effective strategy to improve overall student outcomes.